Videogame pricing data
========================================================

Videogame consumers may wonder what kind of game they should "invest" in.  What kind of games and what kind of condition should they be kept in?



```{r}
price.guide <- read.csv("price-guide.csv",na.strings=c(""))
head(price.guide)

```
First some definitions:  
- ID is just a generic ID for the game
- console name is the name for the console
- product name is the name of the game or accessory
- loose price is the price of the product without case or manual
- cib price is the price of the product with case and manual
- new price is the price of the the new unopened product
- genre is the genre of the product
- release date is the release date of the product

______________________________________
Next we do some cleanup of the data, fixing the dates as well as removing the '$' sign from the prices.


```{r}
price.guide$release.date = as.Date(price.guide$release.date)

price.guide$loose.price = as.numeric(gsub("\\$","",price.guide$loose.price))
price.guide$cib.price = as.numeric(gsub("\\$","",price.guide$cib.price))
price.guide$new.price = as.numeric(gsub("\\$","",price.guide$new.price))
```

We can see from the plots below that the prices are a bit skewed.

```{r fig.width=7, fig.height=6}
plot(price.guide$loose.price)
plot(price.guide$new.price)
summary(price.guide)
```

___________________________________________
Therefore we do the following log transformations

```{r}
price.guide$logloose = log(price.guide$loose.price)
price.guide$logcib = log(price.guide$cib.price)
price.guide$lognew = log(price.guide$new.price)
```
A panel equation that we will be using later
```{r}
p.regression = function(x,y) {
  panel.xyplot(x,y)
  panel.abline(lm(y~x))
}

library(lattice)
```

```{r fig.width=10, fig.height=10}
xyplot(logcib~logloose | console.name,data = price.guide, panel=p.regression)
```
This graph shows, somewhat unsurprisingly, that as loose prices increase, cib prices also increase. 

___________________________________________

```{r fig.width=10, fig.height=10}
xyplot(lognew~release.date | genre,data = price.guide, panel=p.regression)
```
This plot is a bit more interesting.  If we ignore music and controllers, we see that generally the older games have a higher new price than younger ones.  However, this does not seem to be the case for FPS (first person shooter) games
```{r fig.width=10, fig.height=10}
xyplot(lognew~release.date | genre,data = subset(price.guide,genre=='FPS'), panel=p.regression)
xyplot(logloose~release.date | console.name,data = subset(price.guide,genre=='FPS'), panel=p.regression)
```
A reason that newer FPS games seem to enjoy higher new prices could be that FPSs have become much more popular in recent years than in years past.

____________________________________________
Finally lets define a spread of the new price - the loose price
```{r}
price.guide$spread = price.guide$lognew-price.guide$logloose
summary(price.guide$spread)
boxplot(price.guide$spread)
```
So for the most part there is not that much of a spread between the new and loose prices


_______________________________________________


Over time, is there one genre or console where new games are prized much more than loose games?

No, it seems that over time that spread decreases, this makes sense as older games in mint condition are considered more valuable
```{r fig.width=10, fig.height=10}
xyplot(spread~release.date | genre,data = price.guide, panel=p.regression)

xyplot(spread~release.date,data=price.guide, panel=p.regression)

xyplot(spread~release.date | console.name,data = price.guide, panel=p.regression)
```

The Atari 400 on a further look
```{r fig.width=10, fig.height=10}
xyplot(spread~release.date,data = subset(price.guide,console.name=='Atari 400'), panel=p.regression)
```

________________________________________
A conclusion to this is to 'invest' in new FPS games